\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-16 paper.qmd
Removed unnecessary figures

::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/tsne_layout.png)

![](figures/five_gau_clusters/2d_model_tsne.png)

![](figures/five_gau_clusters/sc_tsne_3.png)

The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/RASEE7N5MbM>). 
:::

::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/umap_layout.png)

![](figures/five_gau_clusters/2d_model_umap.png)

![](figures/five_gau_clusters/sc_umap_3.png)

The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/iG4bCPkJilw>). 
:::

::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/pacmap_layout.png)

![](figures/five_gau_clusters/2d_model_pacmap.png)

![](figures/five_gau_clusters/sc_pacmap_2.png)

The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and shows *flat* shaped clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/z07cKXi8EJQ>). 
:::

>>>>

<!-- ::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"} -->
<!-- ![](figures/five_gau_clusters/tsne_layout.png) -->

<!-- ![](figures/five_gau_clusters/2d_model_tsne.png) -->

<!-- ![](figures/five_gau_clusters/sc_tsne_3.png) -->

<!-- The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/RASEE7N5MbM>).  -->
<!-- ::: -->

<!-- ::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"} -->
<!-- ![](figures/five_gau_clusters/umap_layout.png) -->

<!-- ![](figures/five_gau_clusters/2d_model_umap.png) -->

<!-- ![](figures/five_gau_clusters/sc_umap_3.png) -->

<!-- The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/iG4bCPkJilw>).  -->
<!-- ::: -->

<!-- ::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"} -->
<!-- ![](figures/five_gau_clusters/pacmap_layout.png) -->

<!-- ![](figures/five_gau_clusters/2d_model_pacmap.png) -->

<!-- ![](figures/five_gau_clusters/sc_pacmap_2.png) -->

<!-- The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and shows *flat* shaped clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/z07cKXi8EJQ>).  -->
<!-- ::: -->

### 2024-08-16 paper.tex
Removed unnecessary figures

The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five
well-separated clusters which evident these methods effectively preserve
the global structure. In tSNE (Figure~\ref{fig-gau-tsne-sc} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (Figure~\ref{fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
central cluster and the remaining four spread out in different
directions (Figure~\ref{fig-gau-pacmap-sc} (a)).

The tSNE and UMAP shows \emph{filled out} clusters which provide
evidence that these methods preserve the local structure
(Figure~\ref{fig-gau-tsne-sc} (c) and Figure~\ref{fig-gau-umap-sc} (c)).
On the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
and evident that PaCMAP fail to capture the within-cluster variation
(Figure~\ref{fig-gau-pacmap-sc} (c)).

@@

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/tsne_layout.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/2d_model_tsne.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%

\caption{\label{fig-gau-tsne-sc}The tSNE layout, model in \gD{}, and a
view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and tries to
\emph{filled out} the clusters. (The \textbf{langevitour} software is
used to view the data with a tour, and the full video is available at
\url{https://youtu.be/RASEE7N5MbM}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/umap_layout.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/2d_model_umap.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%

\caption{\label{fig-gau-umap-sc}The UMAP layout, model in \gD{}, and a
view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and tries to
\emph{filled out} the clusters, but not as much as tSNE. (The
\textbf{langevitour} software is used to view the data with a tour, and
the full video is available at \url{https://youtu.be/iG4bCPkJilw}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/pacmap_layout.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/2d_model_pacmap.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%

\caption{\label{fig-gau-pacmap-sc}The PaCMAP layout, model in \gD{}, and
a view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and shows
\emph{flat} shaped clusters. (The \textbf{langevitour} software is used
to view the data with a tour, and the full video is available at
\url{https://youtu.be/z07cKXi8EJQ}).}

>>>>

The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five
well-separated clusters which evident these methods effectively preserve
the global structure. In tSNE (\textbf{?@fig-gau-tsne-sc} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (\textbf{?@fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
central cluster and the remaining four spread out in different
directions (\textbf{?@fig-gau-pacmap-sc} (a)).

The tSNE and UMAP shows \emph{filled out} clusters which provide
evidence that these methods preserve the local structure
(\textbf{?@fig-gau-tsne-sc} (c) and \textbf{?@fig-gau-umap-sc} (c)). On
the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
and evident that PaCMAP fail to capture the within-cluster variation
(\textbf{?@fig-gau-pacmap-sc} (c)).

@@
